The letter

### 2010-06-29 ltx/PLoS/LaTeX_Manuscript.tex
added edits from David

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{A Data Driven Framework for Patient-Specific Neural Field Modelling}
}
% Insert Author names, affiliations and corresponding author email.
\\
Dean R. Freestone$^{1,2,3,\ast}$, 
Parham Aram$^{4}$, 
Michael Dewar$^{3,5}$,
Kenneth Scerri$^{6}$,
David B. Grayden$^{1}$,
Visakan Kadirkamanathan$^{4}$
\\
\bf{1} Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia
\\
\bf{2} The Bionic Ear Institute, East Melbourne, VIC, Australia
\\
\bf{3} Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK
\\
\bf{4} Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK
\\
\bf{5} Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA
\\
\bf{6} Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta
\\
$\ast$ E-mail: dfreestone@bionicear.org
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
This paper presents a framework for creating patient-specific neural field models from electrophysiological data. The Wilson and Cowen or Amari style neural field equations are used to form a parametric model, where the parameters are estimated from data. To illustrate the estimation framework, data is generated using the neural field equations incorporating modelled sensors, so a comparison can be made between estimated and true parameters. To facilitate state and parameter estimation, we introduce a method to reduce the continuum neural field model, using a basis function decomposition, forming a finite-dimensional state-space model. Spatial frequency analysis methods are introduced for model selection by systematically specifying the basis function configuration required to capture the dominant characteristics of the neural field. The estimation procedure consists of a two-stage iterative algorithm incorporating the unscented Rauch-Tung-Striebel smoother for state estimation and a least squares algorithm for parameter estimation. The results show that it is theoretically possible to reconstruct the neural field and estimate intracortical connectivity and synaptic dynamics with the proposed framework. The results also illustrate the loss of high spatial frequency information as the cost incurred by the model reduction procedure. This framework provides a link between patient-specific neurophysiological data and theoretical neural fields models. This link may lead to greater understanding of cortical dynamics at the meso and macroscopic scale where diseases such as epilepsy are manifested.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
With recent advances in medical imaging techniques and computing power we are entering a new age of neuroscientific research, where patient-specific information is becoming incredibly detailed. In order to make the best use of this information, patient-specific mathematical models should be formed. This is particularly poignant where considering poorly understood complex systems such as large-scale dynamics of the human brain. There currently exists a library of mathematical models describing large-scale brain dynamics, providing a theoretical explanation of various phenomena associated with various cognitive states and disease. In this paper, we have created a framework for creating a patient specific mathematical model, forming a bridge between theory and electrophysiological data. Within this framework, we propose rules for designing experiments and we derive the mathematical formula to create the model. To demonstrate the utility of this framework, we provide an example using synthetic data that shows a comparison between true and estimated model characteristics. The ability to create patient-specific models of large-scale brain dynamics has the potential to revolutionise the treatment of diseases such as epilepsy, where techniques from engineering may be used to prevent abnormal neural oscillations.

\section*{Introduction}
Generating physiologically plausible neural field models are of great importance for studying brain dynamics at the meso and macroscopic scale. While our understanding of the function of neurons is well developed, the overall behaviour of the brain's meso and macro-scale dynamics remains largely theoretical. Understanding the brain at this level is extremely important since it is at this scale that pathologies such as epilepsy, Parkinson's disease and schizophrenia are manifested.

@@

Unfortunately, the use of these models in the clinic has been limited, since they are constructed for ``general'' brain dynamics whereas pathologies almost always have unique underlying patient-specific causes. Patient-specific data from electrophysiological recordings is readily available in the clinical setting, particularly from epilepsy surgery patients, suggesting an opportunity to make the patient-specific link to models of cortical dynamics. Furthermore, recent technological advances have driven an increased level of sophistication in recording techniques, with dramatic increases in spatial and temporal sampling~\cite{Brinkmann2009}. However, the meso and macroscopic neural dynamic state is not directly observable in neurophysiological data, making predictions of the underlying physiology inherently difficult.

@@

An estimation framework for neural field models known as dynamical causal modelling (DCM) \cite{David2003,David2006} has recently been proposed for studying evoked potential dynamics. Via a Bayesian inference scheme, DCM estimates the long range (cortico-cortical) connectivity structure between the specific isolated brain regions that best explains a given data set using the Jansen and Ritt equations. Another recent publication describing a parameter estimation method with a neural field model used an unscented Kalman filter with modified WC neural field equations~\cite{schiff2008kalman}. This work takes a system theoretic approach to the neural estimation problem, successfully demonstrating that it is possible to perform state estimation on spatio-temporal neural fields. This marks the first step in what has the potential to revolutionise the treatment of many neurological diseases where therapeutic electrical stimulation is viable.

@@

	\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}. 

@@

where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. The connectivity kernel is typically a ``Mexican hat'' function, which describes local excitation and surround inhibition. To demonstrate flexibility in estimation algorithm, a third component of the kernel is introduced describing weak longer-range excitation. An example of the connectivity kernel is shown in in Figure~\ref{fig:2d_kernel}. The exact shape of this kernel is assumed to vary across patients, and hence needs to be inferred from data.

@@

The mapping between the membrane voltage and the electrophysiological data (iEEG/LFP) is modelled using the observation function that incorporates sensors with a spatial extent by
\begin{equation}
    \label{eq:ObservationEquation}
	\mathbf{y}_t =
	\int_{\Omega}{
	    m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
	\textrm{d}\mathbf{r}'} + 
	\boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field, $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$, $\mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$ denotes the multivariate normal distribution with mean zero and covariance matrix $\boldsymbol{\Sigma}_{\varepsilon}$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined by 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)},
\end{equation}
where $\sigma_m$ sets the sensor width.

@@

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. An example of the field decomposition is given in Figure~\ref{fig:FieldDecomposition}.%The choice of Gaussian basis functions can be justified by the existence of the so called bump solutions for this class of model, which have a Gaussian shape~\cite{Coombes2005}.

@@

   	\caption{{\bf Example of a one-dimensional field decomposition.} A continuous one-dimensional field decomposed by a finite number of basis functions scaled by the state vector. The dashed line depicts the field, the black solid lines shows the Gaussian basis functions and the vertical lines show the position and amplitude of the states.} 

@@

Now we define the state disturbance as
\begin{equation}\label{eq:Wt} 
	\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )\textrm{d}\mathbf{r}} 
\end{equation}
which is a zero mean normally distributed white noise process with covariance (see the Supporting Information, S1, for the derivation)
\begin{equation}
	\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
\end{equation}
Under the assumption that the basis function decomposition is accurate, the observation equation of the reduced model is found by substituting equation~\ref{DefFieldDecomp} into~\ref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t\textrm{d}\mathbf{r}'} + \boldsymbol{\varepsilon}_t. 
\end{equation}
The observation equation is linear and can be written in the more compact form
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where the observation matrix is 
\begin{equation}
	\mathbf{C} = \left[
	\begin{array}{{ccc}} 
		c_{1,1} & \dots & c_{1,L} \\
		\vdots & \ddots & \vdots \\
		c_{N,1} & \dots & c_{N,L} 
	\end{array}
	\right] 
\end{equation}
and 
\begin{equation}
	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
\end{equation}
Now we have the final form of the state-space model where
\begin{equation}\label{eq:finalformstatespacemodel}
	\mathbf{x}_{t+1} = Q(\mathbf{x}_t) +\mathbf{e}_t.
\end{equation}
\begin{equation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t
\end{equation}
and 
\begin{equation}\label{eq:QmatrixForSigmapoints}
	Q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t.
\end{equation}

@@

In practice, it is difficult to estimate the bandwidth of the cortex using traditional electrophysiological measurements, possibly preventing us from positioning sensors in accordance to equation~\ref{eq:MinimumSensorDistance}. However, we envisage it may be possible to estimate the spatial bandwidth using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Nevertheless, spectral aliasing can still be avoided by a proper choice of the spatial sampling distance given the sensors' spectral characteristics. The spatial extent of the sensors results in a spectral low-pass action, thus providing spatial anti-aliasing filtering. Such sensors therefore can only sense the field up to a known spatial bandwidth, denoted by $\boldsymbol{\nu}_{cy}$. Aliasing can then be avoided by positioning the sensors according to equation~\ref{eq:MinimumSensorDistance}, replacing $\boldsymbol{\nu}_c$ with $\boldsymbol{\nu}_{cy}$.

@@

\subsection{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section we describe the procedure for estimating the states, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic dynamics, $\xi$. The estimation process is a two part iterative algorithm, consisting of a state estimation step followed by a parameter estimation step. At each iteration, the sequence of estimated state vectors is used to update the parameter set. The resulting parameters are then used when estimating the new state vector sequence for the next iteration. The procedure stops when the parameters converge. The algorithm is initialised using a bounded random state vector sequence guaranteeing that the initial estimated parameter set forms a stable kernel.

The unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{equation}\label{eq:sigmapoints1}
	\mathcal X_{0}=\bar x 
\end{equation}
\begin{equation}
	\mathcal X_{i}=\bar x+\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, L 
\end{equation}
\begin{equation}\label{eq:sigmapoints2}
	\mathcal X_{i}=\bar x-\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_{i- L}, \quad i= L+1, \dots, 2 L 
\end{equation}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for the filtering or smoothing, $\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the weighted matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points, $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$ and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
\begin{equation}
	\mathbf W_0^{(m)}=\frac{\lambda}{ L+\lambda} 
\end{equation}
\begin{equation}
	\mathbf W_0^{(c)}=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta) 
\end{equation}
\begin{equation}
	\mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L, 
\end{equation}
where the superscripts $m$ and $c$ denote mean and covariance. Since the observation equation is linear, (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. The cross-covariance matrix of the states, $\mathbf M$, is also required for computing the smoother gain. A summary of the steps in the URTSS algorithm is given in Table~\ref{tab:UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used.

>>>>

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{A Data Driven Framework for Patient-Specific Neural Field Modelling}
}
% Insert Author names, affiliations and corresponding author email.
\\
Dean R. Freestone$^{1,2,3,\ast}$, 
Parham Aram$^{4}$, 
Michael Dewar$^{3,5}$,
Kenneth Scerri$^{6}$,
David B. Grayden$^{1,2}$,
Visakan Kadirkamanathan$^{4}$
\\
\bf{1} Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia
\\
\bf{2} The Bionic Ear Institute, East Melbourne, VIC, Australia
\\
\bf{3} Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK
\\
\bf{4} Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK
\\
\bf{5} Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA
\\
\bf{6} Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta
\\
$\ast$ E-mail: dfreestone@bionicear.org
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
This paper presents a framework for creating patient-specific neural field models from electrophysiological data. The Wilson and Cowen or Amari style neural field equations are used to form a parametric model, where the parameters are estimated from data. To illustrate the estimation framework, data is generated using the neural field equations incorporating modelled sensors, so a comparison can be made between estimated and true parameters. To facilitate state and parameter estimation, we introduce a method to reduce the continuum neural field model, using a basis function decomposition, forming a finite-dimensional state-space model. Spatial frequency analysis methods are introduced for model selection by systematically specifying the basis function configuration required to capture the dominant characteristics of the neural field. The estimation procedure consists of a two-stage iterative algorithm incorporating the unscented Rauch-Tung-Striebel smoother for state estimation and a least squares algorithm for parameter estimation. The results show that it is theoretically possible to reconstruct the neural field and estimate intracortical connectivity and synaptic dynamics with the proposed framework. The results also illustrate the loss of high spatial frequency information as the cost incurred by the model reduction procedure. This framework provides a link between patient-specific neurophysiological data and theoretical neural field models. This link may lead to greater understanding of cortical dynamics at the mesoscopic and macroscopic scales where diseases such as epilepsy are manifested.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
With recent advances in medical imaging techniques and computing power, we are entering a new age of neuroscientific research, where patient-specific information is becoming incredibly detailed. In order to make the best use of this information, patient-specific mathematical models should be formed. This is particularly poignant where considering poorly understood complex systems such as large-scale dynamics of the human brain. There currently exists a library of mathematical models describing large-scale brain dynamics, providing a theoretical explanation of various phenomena associated with various cognitive states and disease. In this paper, we have created a framework for creating a patient-specific mathematical model, forming a bridge between theory and electrophysiological data. Within this framework, we propose rules for designing experiments and we derive the mathematical formula to create the model. To demonstrate the utility of this framework, we provide an example using synthetic data that shows a comparison between true and estimated model characteristics. The ability to create patient-specific models of large-scale brain dynamics has the potential to revolutionise the treatment of diseases such as epilepsy, where techniques from engineering may be used to prevent abnormal neural oscillations.

\section*{Introduction}
Generating physiologically plausible neural field models are of great importance for studying brain dynamics at the mesoscopic and macroscopic scale. While our understanding of the function of neurons is well developed, the overall behaviour of the brain's mesoscopic and macroscopic scale dynamics remains largely theoretical. Understanding the brain at this level is extremely important since it is at this scale that pathologies such as epilepsy, Parkinson's disease and schizophrenia are manifested.

@@

Unfortunately, the use of these models in the clinic has been limited, since they are constructed for ``general'' brain dynamics whereas pathologies almost always have unique underlying patient-specific causes. Patient-specific data from electrophysiological recordings is readily available in the clinical setting, particularly from epilepsy surgery patients, suggesting an opportunity to make the patient-specific link to models of cortical dynamics. Furthermore, recent technological advances have driven an increased level of sophistication in recording techniques, with dramatic increases in spatial and temporal sampling~\cite{Brinkmann2009}. However, the mesoscopic and macroscopic neural dynamic state is not directly observable in neurophysiological data, making predictions of the underlying physiology inherently difficult.

@@

An estimation framework for neural field models known as dynamical causal modelling (DCM) \cite{David2003,David2006} has recently been proposed for studying evoked potential dynamics. Via a Bayesian inference scheme, DCM estimates the long range (cortico-cortical) connectivity structure between the specific isolated brain regions that best explains a given data set using the Jansen and Ritt equations~\cite{Jansen1995}. Another recent publication describing a parameter estimation method with a neural field model used an unscented Kalman filter with modified WC neural field equations~\cite{schiff2008kalman}. This work takes a system theoretic approach to the neural estimation problem, successfully demonstrating that it is possible to perform state estimation on spatio-temporal neural fields. This marks the first step in what has the potential to revolutionise the treatment of many neurological diseases where therapeutic electrical stimulation is viable.

@@

	\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}, 

@@

where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. The connectivity kernel is typically a ``Mexican hat'' function, which describes local excitation and surround inhibition. An example of the connectivity kernel is shown in in Figure~\ref{fig:2d_kernel}. To demonstrate flexibility in the estimation algorithm, a third component of the kernel is introduced describing weak longer-range excitation. The exact shape of this kernel is assumed to vary across patients, and hence needs to be inferred from data.

@@

The mapping between the membrane voltage and the electrophysiological data (iEEG/LFP) is modelled using the observation function that incorporates sensors with a spatial extent by
\begin{equation}
    \label{eq:ObservationEquation}
	\mathbf{y}_t =
	\int_{\Omega}{
	    m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
	\textrm{d}\mathbf{r}'} + 
	\boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field, $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$. $\mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$ denotes the multivariate normal distribution with mean zero and covariance matrix $\boldsymbol{\Sigma}_{\varepsilon}$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined by 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)},
\end{equation}
where $\sigma_m$ sets the sensor width.

@@

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. An example of a one-dimensional field decomposition is given in Figure~\ref{fig:FieldDecomposition}.%The choice of Gaussian basis functions can be justified by the existence of the so called bump solutions for this class of model, which have a Gaussian shape~\cite{Coombes2005}.

@@

   	\caption{{\bf Example of a one-dimensional field decomposition.} A continuous, one-dimensional field decomposed by a finite number of basis functions scaled by the state vector. The dashed line depicts the field, the black solid lines shows the Gaussian basis functions and the vertical lines show the position and amplitude of the states.} 

@@

Now we define the state disturbance as
\begin{equation}\label{eq:Wt} 
	\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )\textrm{d}\mathbf{r}},
\end{equation}
which is a zero mean normally distributed white noise process with covariance (see the Supporting Information, S1, for the derivation)
\begin{equation}
	\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
\end{equation}
Under the assumption that the basis function decomposition is accurate, the observation equation of the reduced model is found by substituting equation~\ref{DefFieldDecomp} into~\ref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t\textrm{d}\mathbf{r}'} + \boldsymbol{\varepsilon}_t. 
\end{equation}
The observation equation is linear and can be written in the more compact form
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where the observation matrix is 
\begin{equation}
	\mathbf{C} = \left[
	\begin{array}{{ccc}} 
		c_{1,1} & \dots & c_{1,L} \\
		\vdots & \ddots & \vdots \\
		c_{N,1} & \dots & c_{N,L} 
	\end{array}
	\right] 
\end{equation}
and 
\begin{equation}
	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
\end{equation}
Now we have the final form of the state-space model where
\begin{equation}\label{eq:finalformstatespacemodel}
	\mathbf{x}_{t+1} = Q(\mathbf{x}_t) +\mathbf{e}_t
\end{equation}
\begin{equation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t
\end{equation}
and 
\begin{equation}\label{eq:QmatrixForSigmapoints}
	Q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t.
\end{equation}

@@

In practice, it is difficult to estimate the bandwidth of the cortex using traditional electrophysiological measurements, possibly preventing us from positioning sensors in accordance to equation~\ref{eq:MinimumSensorDistance}. However, we envisage it may be possible to estimate the spatial bandwidth using other modalities with higher spatial resolution such as fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Nevertheless, spectral aliasing can still be avoided by a proper choice of the spatial sampling distance given the sensors' spectral characteristics. The spatial extent of the sensors results in a spectral low-pass action, thus providing spatial anti-aliasing filtering. Such sensors, therefore, can only sense the field up to a known spatial bandwidth, denoted by $\boldsymbol{\nu}_{cy}$. Aliasing can then be avoided by positioning the sensors according to equation~\ref{eq:MinimumSensorDistance}, replacing $\boldsymbol{\nu}_c$ with $\boldsymbol{\nu}_{cy}$.

@@

\subsection{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section, we describe the procedure for estimating the states, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic dynamics, $\xi$. The estimation process is a two part iterative algorithm, consisting of a state estimation step followed by a parameter estimation step. At each iteration, the sequence of estimated state vectors is used to update the parameter set. The resulting parameters are then used when estimating the new state vector sequence for the next iteration. The procedure stops when the parameters converge. The algorithm is initialised using a bounded random state vector sequence guaranteeing that the initial estimated parameter set forms a stable kernel.

The unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{align}\label{eq:sigmapoints1}
	\mathcal X_{0}&=\bar x \\
	\mathcal X_{i}&=\bar x+\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, L \\
	\mathcal X_{i}&=\bar x-\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_{i- L}, \quad i= L+1, \dots, 2L 
\end{align}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for the filtering or smoothing, $\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points, $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$ and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
\begin{align}
	\mathbf W_0^{(m)}&=\frac{\lambda}{ L+\lambda} \\
	\mathbf W_0^{(c)}&=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta) \\
	\mathbf W_i^{(m)}&=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L, 
\end{align}
where the superscripts $m$ and $c$ denote mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}), the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. The cross-covariance matrix of the states, $\mathbf M$, is also required for computing the smoother gain. A summary of the steps in the URTSS algorithm is given in Table~\ref{tab:UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$, and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used.

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new estimate of the state vector sequence. To create the least squares parameter estimator, we first define the $L \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from an iteration of the URTSS, we can write
\begin{align*}
	\mathbf x_{1} &= \mathbf{q}(\mathbf x_0) \boldsymbol{\theta}+\xi\mathbf x_0+\mathbf e_0 \\
	\mathbf x_{2} &= \mathbf{q}(\mathbf x_1) \boldsymbol{\theta}+\xi\mathbf x_1+\mathbf e_1  \\
	&\vdots& \\
	\mathbf x_{T}&=\mathbf{q}(\mathbf x_{T-1}) \boldsymbol{\theta}+\xi\mathbf x_{T-1}+\mathbf e_{T-1}. 
\end{align*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathcal W+\mathbf{e}, 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\mathbf x_{1}\\
		\mathbf x_{2}\\
		\vdots\\
		\mathbf x_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\mathbf x_0)& \mathbf x_{0}\\
		\mathbf q(\mathbf x_1)& \mathbf x_{1}\\
		\vdots & \vdots\\
		\mathbf q(\mathbf x_{T-1})& \mathbf x_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathcal W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the LS parameter estimates, $ \mathcal{\hat{W}}$, are
\begin{equation}
	\mathcal{\hat{W}}=(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}

% Results and Discussion can be combined.
\section*{Results}\label{ResultsSection} In this section, we provide a demonstration of the estimation procedure using the new methods that are described above. The full neural field model described in Section~\ref{NeuralModelSection} was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Parameters}. Note, the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is due to the fact that the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the space step used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field can be seen in Figure~\ref{fig:experimental design}(A). Figure~\ref{fig:experimental design}(B) shows an example of data from the modelled sensors and Figure~\ref{fig:experimental design}(C) shows the power spectral density (PSD) of the simulated time-series. The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.
